Flaring star-planet interactions in star-planet systems in Kepler and TESS

### 2023-03-23 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
\begin{abstract}
    The star-planet systems discovered so far are nothing like the Solar System. Many stars host planets in orbits much closer than Mercury, where it takes them mere days or even hours to complete a full revolution. In some of these systems, the planets are so close that they plow through the extended magnetosphere of the star, and are expected to trigger flares on their surfaces. So far, however, these flaring star-planet interactions escape our detection methods. If found, the planets and the flare they cause become localized probes of the parts of the stellar magnetosphere that extends from the location of one to the other. A measurement of these interactions constrains mass loss from the star, and potentially also the magnetic field of the planet.

    In this work, we searched star-planet systems observed by Kepler and TESS for flaring star-planet interactions. We compiled a catalog of all flaring star-planet systems known to date, and characterized the detected flares detected. We then searched for flares in excess of the intrinsic flaring of the star, i.e., flares triggered by the interaction with the planet in phase with its orbit.

    We find XXXX star-planet systems with > XXXX flares. For XXXX of them, we can quantify both the intensity of excess flaring, and the expected amplitude of interaction based on the system's magnetic field strength, orbital parameters, rotation and planet radius. Most systems are neither expected to show interactions, nor do they show any excess flaring. Among those systems that have high expected power of interaction, we see two branches. An inactive one, where no signs of excess flaring appear, and an active one %We find that the trend of higher interaction signal agrees with the expected strength, but also shows signs of intermittency, both for individual systems, and in comparison between ones with similar expected amplitude. HIP 67522, a 17 Myr young Sun with a Jupiter in a 9 day orbit, shows both one of the highest expected and the highest measured interaction amplitudes, albeit only at the $2.5\sigma$ level. 
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. In particular, the past two decades revealed that planets can occur in orbits so close to their host star that they could be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, gravitational and magnetic, both short- and long-term, on its host -- a phenomenon unknown to the Solar System. The individual processes are typically subsumed under the term star-planet interactions that emphasizes that they are bidirectional, affecting both planet, and star.

Understanding short-term, magnetic star-planet interactions holds promise to provide important insights into the habitability of the entire system. Planets involved in such interactions can be viewed as in-situ probes of the system's space weather, and could prospectively reveal otherwise difficult to measure planetary properties such as the strength and extent of their magnetic fields.

Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star longer than it takes for the effects of the planet to travel to the surface of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceed the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in some way. \citet{lanza2012starplanet, lanza2018closeby} and \citet{zarka2007plasma, saur2013magnetic} suggest different mechanisms for when and how the energy dissipation takes place, and how much energy can be transferred and dissipated at all.

One way of dissipating the energy in the magnetic field is through flares. 
Flares are explosions in stellar atmospheres that occur in the stellar corona as a way to dissipate energy stored in the magnetic field by the motion of footpoints of coronal loops. They are strong and impulsive eruptions released by reconnection and subsequent relaxation of field lines~\citep{svestka1976solar,priest2002magnetic}. We can observe flares in nearly all electromagnetic bands -- from radio to X-ray~\citep{benz2010physical}. In contrast to solar flares, stellar flares often are extremely energetic, sometimes enhancing the stellar flux by orders of magnitude~\citep{maehara2012superflares, shibayama2013superflares, paudel2018k2}. But since they occur randomly in time, individual events can be hard to catch.

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data suitable for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly expanding the treasury~\citep{gunther2020stellar}, which will grow even further after the launch of the PLATO mission in 2026~\citep{rauer2014plato}. As will become clear in the following paragraphs, the key advantage of these missions is their quasi uninterrupted monitoring of star-planet systems that covers all phases of their planets' orbits.

The absolute power of flaring star-planet interactions is difficult to constrain, so we do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand that power is an upper limit, so the observed energies might be lower. The planet might for instance be triggering flares prematurely, leading an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism, so it might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

@@

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's architecture and magnetic situation, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable architectures and magnetic field states, the footpoint visibility will be modulated in phase with the planetary orbit. The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\citet{fischer2019timevariable} and \citet{howard2021evryflare}

@@

\subsection{Rotation periods}
\label{sec:data:rotationperiods}
We adopt available rotation period values from the literature. Almost all stem from light curve variability~\citep{angus2018inferring, mazeh2015photometric, mcquillan2013stellar, mcquillan2014rotation, luger2017sevenplanet, stock2020carmenes, deleon202137, torres2017validation, stefansson2020habitable, zicher2022one, ment2021toi, rizzuto2020tess}, and only one from periodic variation in chromospheric lines~\citep{demangeon2021warm}. When uncertainties are not given, we conservatively assume $10\%$ uncertainty, WHERE ARE THE PERIODS FROM.

@@

The estimate is consistent with existing Zeeman broadening for, e.g. AU Mic (3010 G~\cite{reiners2022magnetism}).
\subsection{Power of star-planet interaction}
\label{sec:methods:pspi}
We consider two theories for the mechanism behind flaring star-planet interactions, the stretch-and-break and the Alfv\'en wing mechanisms. For both of them, we include the case of a magnetized and an unmagnetized planet, for a total of four estimates of the expected power of magnetic star-planet interaction.

@@

\subsection{Eccentricity}
\label{sec:discussion:eccentricity}
If the planetary orbit is highly eccentric, the difference between the semi-major and semi-minor axis can be significantly larger than the variability of the Alfv\'en radius. If then apoastron is outside this radius, and periastron within, the planet may find itself in the sub-Alfv\'enic zone only for a short amount of time, namely during the planet's rapid periastron passage. This was previously exploited in the case of HD 189733, and also for binary interaction in the V 773 Tau system. In our sample, the estimates for eccentricities of the innermost planets are all moderate to low, with $e<XXXX$. The exception is K2-25, with a moderate $e=0.45$, which may contribute to the increased contrast of excess flares. All else equal, closer proximity increases the strength of interaction. Perhaps, K2-25 even moves in and out of the sub-Alfv\'enic zone between periastron and apoastron, turning interaction on and off every single orbit. Otherwise, there are therefore no planets where we would expect a strongly preferred orbital phase for interaction.

\subsection{Rotational and synodic variability}
\label{sec:discussion:rotsyn}
\citet{fischer2019timevariable} argue that the synodic period, at which the sub-planetary point crosses the same (magnetic) surface element on the star, may result in a clearer interaction signal, because a magnetically active region that is prone to planet-induced flares, will be passed only once per period. However, this assumes that the landscape of active regions on the star is stable throughout the entire observing time, and that the number of such regions in the footpoint passageway is low enough to produce a periodic modulation.

In our sample, the observing baseline is multiple years long for many systems. Even the most rigidly rotating late M dwarfs show variability in their spot modulation on such long timescales. Even if differential rotation was absent and active regions were extremely stable, we will not be able to observe the passage of the sub-planetary point over the flare-prone region every synodic period. For a fraction of the time, this will happen on the back of the star relative to the observer. The true period will therefore be additionally modulated by the visibility, resulting in an intermittency of the signal. As a consequence, to detect a modulation with the synodic period, a significantly longer observing baseline is needed than to detect a modulation with the orbital period. Additionally, if there are multiple active regions that produce flares when the sub-planetary footpoint passes across, the effect with the synodic period is further diluted.

In summary: Even for the few stars, for which the rotational period is known precisely enough~(see Fig.~\ref{fig:coherence_hist}), we need to make assumptions about the number of active regions crossed by the footpoint of interaction, differential rotation, spot evolution, and a significant correlation between the location of the planet and the location of the planet-induced flare on the stellar surface. For modulation with the orbital period, we require only that last correlation between footpoint location and orbital phase. Any realistic number of active regions (one or more), a wide range of degrees of differential rotation (from rigid rotator to solar-type), and half-life duration of active regions (day to years) are acceptable, because the observed effect depends predominantly on the visibility of the footpoint, which is governed by orbital motion. And except for extreme circumstances, the latter will be much more stable over long periods of time than the intrinsically dynamic magnetic field configuration.

\subsection{Intermittency}
\label{sec:discussion:intermittency}
Magnetic star-planet interactions in a given system may sometimes be turned off for episodes of time, and turned on for others. HD 179949 is the prototype example, wherein the orbital modulation of chromospheric indicators was observable for XXXX out of XXXXX epochs~\citep{shkolnik2003evidence,shkolnik2008nature}. It is not clear how long a continuous epoch of interaction or non-interaction might be. This leaves room for many different explanations. Long-term monitoring is needed to investigate the reasons for the temporary cessation of interaction. Candidate mechanisms include: The Alfv\'en radius moving in- and outside the planet's orbit, either over the course of an activity cycle, or through short-term variations in the magnetic field and wind properties caused by, e.g., coronal mass ejections; visibility of the interacting footpoint changing as the magnetic field evolves; the active latitude moving away from the magnetic footpoints passageway, the available magnetic energy decreasing such that the flares or other magnetic interaction indicators still occur but fall below the detection threshold for a period of time.

If the magnetic field of the planet changes or disappears, this might also reduce the intensity of interaction below the detection threshold.

The location of active regions within the passageway of the magnetic footpoints may be such that the interaction occurs only when the region faces away from the observer. This could span multiple orbits of the planet, particularly when the surface is populated by few stable active regions, and the orbital and rotational periods are similar. The flip-flop behavior of active longitudes could then cause a relatively sudden quenching of interaction by moving the interacting region out of view for a period of time.

In our results in Fig.~\ref{fig:adtest_bp}, we can interpret the two branches emerging at high expected power of interaction as due to the intermittency of the phenomenon. TAP 26 and KOI-12 do not show any sign of excess flaring in our data, despite similar expected powers as in HIP 67522 with its $>2\sigma$ signal. In Sections~\ref{sec:results:individualstars:tap26} and \ref{sec:results:individualstars:koi12}, the absence of interaction in TAP 26 and KOI-12 could be explained by an uncertainty in orbital period that smears out the signal, and a super-Alfv\'enic orbit, respectively. Yet intermittency cannot be ruled out. The case HIP 67522 may be more informative. In our data, we combine the observation from two TESS Sectors that are almost 2 years apart. The four phase-correlated flares all occur in the later Sector, while the earlier one contains only two flares, both of them at orbital phases different from each other and the four phase-aligned flares. This can be interpreted as a tentative sign of intermittency of flaring star-planet interactions in this system.

@@

Most of our systems are transiting, and for those we can expect this scenario to be a minor problem. Of the systems that have high expected power but do not show any, only TAP 26 is a planet detected in radial velocity, and is expected to have a relatively high orbital inclination of about $55\pm10$ deg inferred from Zeeman Doppler measurements~\citep{yu2017hot}. The ZD map reveals magnetic field concentrations close to the rotational pole. If the orbital plane is aligned with the stellar rotation axis, the absence of phase-correlated flaring in TAP 26 could be explained by the footpoint of interaction being always in view and therefore never modulated.

\subsection{(Mis-)alignment between orbital, rotational and stellar magnetic axis}
\label{sec:discussion:misalignment}
Another relevant factor is the geometry of the system. If we assume that the magnetic dipole axis and the rotation axis of the star are relatively well-aligned, the spin-orbit alignment may have an influence on interaction strength.

or more broadly system architecture including misalignment of the magnetic axis. See HAT-P-11, or instance, and the many polar orbits in \cite{bourrier2023dream}

\cite{miskovetz2022resolving}  found background stars that were not co-moving HAT-P-11. HAT-P-11 rotates nearly pole-on, and HAT-P-11 b is transiting in a near-polar orbit~\citep{bourrier2023dream}. This may be the reason for absent SPI.

HD 179949 has a strong misalignment of about 70 deg between rotational and magnetic axis. Time-resolved X-ray observations~\citep{acharya2022xray} suggest that the magnetic interaction footpoint is close to the stellar rotational pole, which would be close to the magnetic equator of the strongly dipolar field of HD 179949

\subsection{Spin-orbit commensurability}
\label{sec:discussion:commensurability}
In Fig.~\ref{fig:spinorbit}, we split our sample into three categories. The first contains the systems where the expected interaction is lower than that in TOI-540. The second includes the systems where a high power of interaction is expected, but none is measured (e.g., TAP 26, KOI-12). The final third category contains all system where the data suggest a correlation between expected power of interaction and measured flaring star-planet interaction signal (e.g., HIP 67522, AU Mic). We search for spin-orbit commensurability with integers $<10$ in each system, and plot the relative difference between the closest ratio and the actual spin-orbit ratio. Systems with low expected power of interaction span both very close spin-orbit commensurability to none. 
Systems with high expected power are divided: Stars that show a correlation between excess flaring and expected power tend to have closer spin-orbit commensurability than systems that are far away from any low-integer ratio between rotational and orbital period.

While chance alignment is possible, the division invokes the idea that spin-orbit commensurability and magnetic star-planet interaction are related. Either something about our observing method favors the measurement of excess flaring in systems with close spin-orbit commensurability, or there is physical causation at play. In the latter case, the spin-orbit commensurability could be either an effect, or a cause, of measured magnetic star-planet interaction.

@@

\citet{szabo2021changing} note that many systems within the sparsely populated region of close-in planets around rapid rotators~\citep{mcquillan2013stellar} seem close to low-integer spin-orbit resonance. The systems with consistently high expected power of interaction (TAP 26, TOI-540, KOI-12, HIP 67522, K2-25, and AU Mic) are all found within that region, even if KOI-12 and AU Mic lie rather close to its edge. Of these six systems, those that show signs of star-planet interaction, are also in close-spin orbit resonance. We may speculate, that at least some of the mechanisms that explain the dearth of close-in planets around rapidly rotating stars also explain the presence of star-planet magnetic interaction in those systems that are still found there. One may think that planets that remain in the desert are held there by being trapped in the resonance, while for the others, the magnetic torque cause rapid migration out of this region.

>>>>

% Abstract with filler text
\begin{abstract}
    In many of the star-planet systems discovered so far, most planets orbit far within Mercury's orbit, and many are so close to their host that they plow through its extended magnetosphere. In these systems, planets become in-situ probes of the dynamic magnetic fields in their orbits, able to trigger flares on the stellar surface. Until recently, however, these flaring star-planet interactions suffered from small sample sizes, and poor orbital phase coverage.

    In this work, we used the excellent phase coverage of Kepler and TESS to search their entire archives for star-planet systems with flaring star-planet interactions. We compiled a catalog of all star-planet systems with flares, and searched for planet-induced flares, i.e., flares clustering in phase with the planet's orbit.

    Out of 1811 searched systems, 26 are single star systems that show three or more flares each. For these, we quantified the intensity of excess flaring, and compared it against the expected amplitude of interaction based on the system's parameters. Most systems are neither expected to interact, nor do they show excess flaring. Among those systems that have high expected power of interaction, we see two branches. An inactive one, where no signs of excess flaring appear regardless of favorable conditions, and an active one where expected power and excess flaring are correlated. While this trend is tentative, we also found that the two branches form two distinct populations in spin-orbit commensurability. The active branch is closer to low-integer ratios between rotational and orbital period than the inactive branch. While viewing geometry and intermittency of interaction remain possible explanations, spin-orbit resonances hint at a physical link worth exploring to better understand the underlying mechanisms of star-planet interactions.
    %magnetic field strength (using $Ro$-$B$ relations), orbital parameters, and planet radius.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. In particular, the past two decades revealed that planets can occur in orbits so close to their host star that they could be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions that emphasizes that they are bidirectional, affecting both planet, and star.

On the one hand, a close-in planet that changes the star's rotation and activity makes it more difficult for us to model the system's evolution, as it affects fundamental stellar parameters. Inferring stellar age from rotation may be biased, if the star is spun up or spun down relative to stars without close-in companions~(XXXX). Analogously, altered activity levels may mimic an older or younger host than it actually is~\citep{ilic2022tidal}.

On the other hand, understanding short-term, magnetic star-planet interactions -- the focus of this work -- holds promise to provide important insights into the habitability of the entire system~(XXXX). Planets involved in such interactions are in-situ probes of the system's space weather~(XXXX), and could prospectively reveal otherwise difficult to measure planetary properties such as the strength and extent of their magnetic field~\citep{cauley2019magnetic}.

\subsection{Magnetic star-planet interactions}

Magnetic star-planet interactions can only occur if the planet is orbiting within the Alfv\'en radius of the star. It does not need to be there all the time, but for a time period that is longer than it takes for the effects of the planet to travel to the surface of the star. The Alfv\'en radius is the radius at which the stellar wind velocity, which increases with increasing distance from the star, exceeds the Alfv\'en velocity of the magnetized plasma. Beyond this radius, this plasma is disconnected from the star. Alfv\'en waves set off by the planet within the magnetic field they cross cannot propagate back to the star if the wind carries the plasma away faster than these waves can travel. In the sub-Alfv\'enic regime, the wave can reach the star, and deposit its energy in some way in its atmosphere. \citet{lanza2012starplanet, lanza2018closeby} and \citet{zarka2007plasma, saur2013magnetic} suggest different mechanisms for when and how the energy transport and dissipation takes place, and how much energy can be transferred and dissipated at all.

\subsection{Searching for global changes in actvity}

Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but just a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible,poppenhaeger2011correlation, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces a bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability.

\subsection{Flares}

In this study, we will look at flares as another way of dissipating the energy deposited in the magnetic field. Flares are explosions that occur in the stellar corona as a way to dissipate energy stored in the magnetic field by the motion of footpoints of coronal loops. They are strong and impulsive eruptions released by reconnection and subsequent relaxation of field lines~\citep{svestka1976solar,priest2002magnetic}. In contrast to solar flares, stellar flares often are extremely energetic, sometimes enhancing the stellar flux by orders of magnitude~\citep{maehara2012superflares, shibayama2013superflares, paudel2018k2} so that they can be detected in stars up to hundreds of parsecs away~(XXXX). But since they occur randomly in time, individual events can be hard to catch. %We can observe flares in nearly all electromagnetic bands -- from radio to X-ray~\citep{benz2010physical}.

In the 2010s, a wealth of flare observations from space missions like Kepler~\citep{borucki2010kepler} and K2~\citep{howell2014k2} have been catalogued~\citep{davenport2016kepler, paudel2018k2, ilin2021flares}. We now have access to long-term stellar monitoring data available for flare hunting. Currently, the Transiting Exoplanet Survey Satellite~\citep{ricker2015transiting} is rapidly growing the archive~\citep{gunther2020stellar}. The key advantage of these missions for this work is their quasi uninterrupted monitoring of star-planet systems that covers all phases of their planets' orbits, as will become clear in the following.

\subsection{Flaring star-planet interactions}

The absolute power of flaring star-planet interactions is difficult to constrain, so we do not know how energetic planet-induced flares can be. The estimates of the maximum dissipated power~\citep{lanza2018closeby} are well within the regime of superflares typically observed by Kepler and TESS. However, on the one hand, that power is an upper limit, so the observed energies might be lower. The planet might for instance be triggering flares prematurely, leading to an overabundance of low energy flares~\citep{loyd2023flares}. On the other hand, the energy need not be dissipated continuously, particularly in the stretch-and-break mechanism, so it might be higher. We also do not know whether flares triggered by interactions with a companion would have different properties than intrinsic flares. We assume that, since they would be caused by magnetic reconnection triggered by the planet in the stellar corona, they should follow the same process as intrinsic flares.

@@

If we picture the magnetic field lines connecting the planet to the star in the sub-Alfv\'enic zone, we expect that the stellar footpoint of these lines will move along with the planet. Flares triggered by the planet will occur at a preferred orbital phase, namely whenever the footpoint is visible to the observer. Depending on the system's viewing geometry, i.e., the inclination of the rotation axis, the spin-orbit (mis-)alignment, and the extent and configuration of the large scale magnetic field -- the footpoint could be always in view, or never. In expectation, however, in star-planet systems with favorable architectures and magnetic field states, the footpoint visibility will be modulated in phase with the planetary orbit. The resulting phase preference of planet-induced flares introduces a deviation from a random distribution of flares. Space missions like Kepler and TESS provide light curves that cover all phases of the planet's orbit, so that phase-correlated planet-induced flares can be detected against a background of uncorrelated intrinsic flaring.

\subsection{Searching for local changes in activity}

Phase correlated activity, flaring and otherwise, has been searched for in individual systems. After eight years of monitoring of HD 189733, an X-ray active Hot Jupiter host, the observed slight clustering of flaring in orbital phase remains tentative~\citep{pillitteri2022xray}. \citet{maggio2015coordinated} observed a flare-up in X-ray when the Jupiter-sized companion was in periastron on its highly eccentric orbit around its host HD 17156. However, a similar study using chromospheric indicators in the eccentric system HD 80606 found no variation between periastron and apoastron~\citep{figueira2016activity}. In the above cases, non-detection can at least partly be attributed to poor orbital phase coverage, which also limits the significance of any positive detection. Realizing this, \citet{fischer2019timevariable} used K2 data with excellent phase coverage of the orbits of the inner TRAPPIST-1 planets to look for clustered flaring in phase, but found none. Similarly, in \citet{ilin2022searching}, we found hints of excess flaring in the TESS light curves of the young AU Mic system. This mixed picture of individual object studies calls for a more comprehensive and systematic approach.

@@

\subsection{Rotation periods}
\label{sec:data:rotationperiods}
We adopt available rotation period values from the literature. Almost all stem from light curve variability~\citep{angus2018inferring, mazeh2015photometric, mcquillan2013stellar, mcquillan2014rotation, luger2017sevenplanet, stock2020carmenes, deleon202137, torres2017validation, stefansson2020habitable, zicher2022one, ment2021toi, rizzuto2020tess, gunther2020stellar}, and only one from periodic variation in chromospheric lines~\citep{demangeon2021warm}. When uncertainties are not given, we conservatively assume $10\%$ uncertainty. Uncertainties are missing usually when periods were detected using Lomb-Scargle~\citep{lomb1976leastsquares, scargle1982studies}, or similar, periodogram peaks only~\citep{gunther2020stellar, kiraga2007agerotationactivity, grankin2013magnetically, burt2014lickcarnegie}, but also when the measurement is indirect using activity-rotation relations~(only in the case of GJ 3323).

GJ 3082 periods were measured and are consistent both in TESS and in KELT light curves~\citep{gunther2020stellar} to below 1\,s precision. GJ 674 periods are consistent between light curve periodograms~\citep{kiraga2007agerotationactivity} and activity-rotation relations~\citep{boisse2011disentangling} at a $<10\%$ level. GJ 3323 only has measured rotation from activity measurements~\citep{astudillo-defru2017magnetic}, without any periodicities detected in the activity indicators themselves~\citep{astudillo-defru2017harps}. The rotation period for Proxima Cen is very slow, but consistent between different datasets~\citep{anglada-escude2016terrestrial, kiraga2007agerotationactivity}. In general, among the stars without given uncertainties on the rotation period, the period is either 
\begin{itemize}
    \item short, leaving a clear peak in the periodogram, and little doubt about the period after ruling out aliases with $P/2$, as with GJ 3082, or
    \item longer than the Sun's, that is, in the regime where rotation-activity relations are reliable because the stars are in the unsaturated activity regime, as in the case of GJ 674.
\end{itemize}

@@

The estimate is consistent with existing Zeeman broadening for, e.g. AU~Mic (3010 G~\cite{reiners2022magnetism}).
\subsection{Power of star-planet interaction}
\label{sec:methods:pspi}
We consider two theories for the mechanism behind flaring star-planet interactions, the stretch-and-break and the Alfv\'en wing mechanisms. For both of them, we include the case of a magnetized and an unmagnetized planet, for a total of four estimates of the expected power of magnetic star-planet interaction.

@@

\subsection{Eccentricity}
\label{sec:discussion:eccentricity}
If the planetary orbit is highly eccentric, the difference between periastron and apoastron can be significantly larger than the variability of the Alfv\'en radius. If then apoastron is outside this radius, and periastron within, the planet may find itself in the sub-Alfv\'enic zone only for a short amount of time, namely during the planet's rapid periastron passage. This was previously exploited in the case of HD 189733~(XXXX), and also for binary interaction in the V 773 Tau system~(XXXX). In our sample, the estimates for eccentricities of the innermost planets are all moderate to low, with $e\leq 0.35$. The exception is K2-25, with a relative high $e=0.45$, which may contribute to the increased contrast of excess flares. All else equal, closer proximity increases the strength of interaction. Perhaps, K2-25 even moves in and out of the sub-Alfv\'enic zone between periastron and apoastron, turning interaction on and off every single orbit. If periastron passage occurs when the planet is in front of the star relative to the observer, the visibility effect of the interaction footpoint moving in and out of view as the planet orbits is further enhanced by the narrow phase range of the periastron passage.

\subsection{Rotational and synodic variability}
\label{sec:discussion:rotsyn}
Our analysis is the most comprehensive search for flaring star-planet interactions to date, using the full Kepler and most recent TESS archives of light curves. We considered exploiting it beyond the search for variations with orbital period, and additionally look for variability with rotation period, and synodic period. Here, we explain why we ultimately decided not to do so.

Even for the few stars, for which the rotational period is known precisely enough~(see Fig.~\ref{fig:coherence_hist}) for our analysis, we need to make a large number of assumptions. These include the number of active regions crossed by the footpoint of interaction, differential rotation, spot evolution, and a significant correlation between the location of the planet and the location of the planet-induced flare on the stellar surface. For modulation with the orbital period, we require only that last correlation between footpoint location and orbital phase. Any realistic number of active regions (one or more), a wide range of degrees of differential rotation (from rigid rotator to solar-type), and half-life duration of active regions (day to years) are acceptable, because the observed effect depends predominantly on the visibility of the footpoint, which is governed by orbital motion. The latter is much more stable over long periods of time than the intrinsically dynamic magnetic field configuration.

\citet{fischer2019timevariable} argue that the synodic period, at which the sub-planetary point crosses the same (magnetic) surface element on the star, may result in a clearer interaction signal, because a magnetically active region that is prone to planet-induced flares, will be passed only once per period. However, this assumes that the landscape of active regions on the star is stable throughout the entire observing time, and that the number of such regions in the footpoint passageway is low enough to produce a periodic modulation. Note that this is a different effect from the geometrical modulation of visibility we look for in modulation with the orbital period in this study.

In our sample, the observing baseline is multiple years long for many systems. Even the most rigidly rotating late M dwarfs show variability in their spot modulation on such long timescales. Even if differential rotation was absent and active regions were extremely stable, we will not be able to observe the passage of the sub-planetary point over the flare-prone region every synodic period. If the planet's rotational axis is inclined, then the passage will regularly occur on the back of the star relative to the observer. The true period will therefore be additionally modulated by the visibility, resulting in an intermittency of the signal. As a consequence, to detect a modulation with the synodic period, a significantly longer observing baseline is needed than to detect a modulation with the orbital period. Additionally, if there are many active regions that produce flares when the sub-planetary footpoint passes across, the effect with the synodic period is further diluted.

\subsection{Intermittency}
\label{sec:discussion:intermittency}
Magnetic star-planet interaction power in a given system may sometimes be low for episodes of time, and high for others. HD 179949 is the prototype example for this phenomenon, wherein the orbital modulation of chromospheric indicators was observable for XXXX out of XXXXX epochs on XXXX years baseline~\citep{shkolnik2003evidence,shkolnik2008nature}. It is not clear how long a continuous epoch of interaction or non-interaction might be. This leaves room for many different explanations. Long-term monitoring is needed to investigate the reasons for the temporary cessation of interaction. Candidate mechanisms for the star include: 
\begin{itemize}
    \item the Alfv\'en radius moving in- and outside the planet's orbit, either over the course of an activity cycle, or through short-term variations in the magnetic field and wind properties caused by, e.g., coronal mass ejections;
    \item visibility of the interacting footpoint changing as the magnetic field evolves;
    \item the active latitude moving away from the magnetic footpoints passageway; or
    \item the available magnetic energy decreasing such that the flares or other magnetic interaction indicators still occur but fall below the detection threshold for a period of time.
\end{itemize}

If the magnetic field of the planet changes or disappears as well, this might also reduce the intensity of interaction below the detection threshold.

The location of active regions within the passageway of the magnetic footpoints may be such that the interaction occurs only when the region faces away from the observer. This could span multiple orbits of the planet, particularly when the surface is populated by few stable active regions, and the orbital and rotational periods are similar. The flip-flop behavior of active longitudes seen in young solar-like stars~(XXXX) could then cause a relatively sudden quenching of interaction by moving the interacting region out of view for a period of time.

In our results in Fig.~\ref{fig:adtest_bp}, we can interpret the two branches emerging at high expected power of interaction as due to intermittency. TAP 26 and KOI-12 do not show any sign of excess flaring in our data, despite similar expected powers as in HIP 67522 with its $2-3\sigma$ signal. In Sections~\ref{sec:results:individualstars:tap26} and \ref{sec:results:individualstars:koi12}, the absence of interaction in TAP 26 and KOI-12 could be explained by an uncertainty in orbital period that smears out the signal, and a super-Alfv\'enic orbit, respectively. Yet intermittency cannot be ruled out. The case HIP~67522 may be more informative. In our data, we combine the observation from two TESS Sectors that are almost 2 years apart. The four phase-correlated flares all occur in the later Sector, while the earlier one contains only two flares, both of them at orbital phases different both from each other and the four phase-aligned flares~(see Fig.~\ref{fig:cumdist_transiting}). This can be interpreted as a tentative sign of intermittency of flaring star-planet interactions in this system.

@@

Most of our systems are transiting, and for those we can expect this scenario to be a minor problem. Of the systems that have high expected power but do not show any, only TAP 26 is a planet detected in radial velocity, and is expected to have a relatively high orbital inclination of about $55\pm10$ deg inferred from Zeeman Doppler measurements~\citep{yu2017hot}. The ZD map reveals magnetic field concentrations close to the rotational pole, and the brightness maps also show a dark spot in the same location. If the orbital plane is aligned with the stellar rotation axis, the absence of phase-correlated flaring in TAP 26 could be explained by the footpoint of interaction being always in view and therefore never modulated.

%If we assume that the magnetic dipole axis and the rotation axis of the star are relatively well-aligned, the spin-orbit alignment may have an influence on interaction strength. For instance,  HAT-P-11 rotates nearly pole-on, and HAT-P-11 b is transiting in a near-polar orbit~\citep{bourrier2023dream}.

% \subsection{(Mis-)alignment between orbital, rotational and stellar magnetic axis}
% \label{sec:discussion:misalignment}

%  Relaxing the assumption of alignment between rotational and magnetic dipole axis, we introduce a further complication that is not only due to viewing geometry.

% or more broadly system architecture including misalignment of the magnetic axis. See HAT-P-11, or instance, and the many polar orbits in \cite{bourrier2023dream}

% \cite{miskovetz2022resolving}  found background stars that were not co-moving HAT-P-11.

% HD 179949 has a strong misalignment of about 70 deg between rotational and magnetic axis. Time-resolved X-ray observations~\citep{acharya2022xray} suggest that the magnetic interaction footpoint is close to the stellar rotational pole, which would be close to the magnetic equator of the strongly dipolar field of HD 179949

\subsection{Spin-orbit commensurability}
\label{sec:discussion:commensurability}
Curiously, in our analysis, systems with high expected power are divided: Stars that show a correlation between excess flaring and expected power tend to have closer spin-orbit commensurability than systems that are far away from any low-integer ratio between rotational and orbital period.

In Fig.~\ref{fig:spinorbit}, we split our sample into three categories. The first contains the systems where the expected interaction is lower than that in TOI-540. The second includes the systems where a high power of interaction is expected, but none is measured (e.g., TAP 26, KOI-12). The final third category contains all system where the data suggest a correlation between expected power of interaction and measured flaring star-planet interaction signal (i.e., K2-25, TOI-540, HIP 67522, and AU Mic). We search for spin-orbit commensurability with integers $<10$ in each system, and plot the relative difference between the closest ratio and the actual spin-orbit ratio. Systems with low expected power of interaction span both very close spin-orbit commensurability to none. However, the second and third group form two separate populations. Non-interacting systems are further away from low-integer commensurability, than interacting ones, even though both have high expected power.

While chance alignment is possible, the division suggests that spin-orbit commensurability and magnetic star-planet interaction are related. Either something about our observing method favors the measurement of excess flaring in systems with close spin-orbit commensurability, or there is physical causation at play. In the latter case, the spin-orbit commensurability could be either an effect, or a cause, of measured magnetic star-planet interaction.

@@

\citet{szabo2021changing} conjecture that many systems within the sparsely populated region of close-in planets around rapid rotators~\citep{mcquillan2013stellar} seem close to low-integer spin-orbit resonance. The systems with consistently high expected power of interaction (TAP 26, TOI-540, KOI-12, HIP 67522, K2-25, and AU Mic) are all found within that region, even if KOI-12 and AU Mic lie rather close to its edge. Of these six systems, those that show signs of star-planet interaction, are also in close-spin orbit resonance. We may speculate, that at least some of the mechanisms that explain the dearth of close-in planets around rapidly rotating stars also explain the presence of star-planet magnetic interaction in those systems that are still found there. One may think that planets that remain in the desert are held there by being trapped in the resonance, while for the others, the magnetic torque causes rapid evolution out of this region through spin-down of the star throug both magnetic braking and star-planet interaction, and the planet's migration to higher orbital periods through the latter mechanism.
